\chapter{Introduction}
\label{chap:introduction}

\section{Problem and Motivation}

Social Networks such as Facebook\footnote{\url{http://www.facebook.com}},
MySpace\footnote{\url{http://www.myspace.com}} or
Twitter\footnote{\url{http://www.twitter.com}} gained millions of members on
their platforms over the last years. They are widely used in both private and
business networking. Also, they allow individuals to present their own user
profile and to share and maintain relationships with other members in the
network. There is no doubt that social networks are a useful instrument and
pose a communication platform which many people are currently using and even
more will be using in the near future~\cite{whitworth2009}.

The nature of social networks is about sharing data but frequently users are as
well presenting their information to an unknown number of strangers. This does
not only raise the question of how useful a social network is for a single
user, but also if there are any drawbacks or risks for an individual or a
company. Users often put a lot of sensitive data into their
profiles, as for example Brown et al. show~\cite{brown2008}. This data however
is stored centrally at the company which provides the network and so the user
looses the control over his or her data~\cite{fraunhofer2008}.

Of course there is a relation between the data which the user entered and the
actual picture of the user. Following that thought, another person is not only
able to extract the data but can obtain new information or interpretations
which were not entered by the actual user.

Many studies such as~\cite{fraunhofer2008,gross2005} show that this can be a
massive intervention into one's privacy. This work however wants to examine
whether such information is a security risk and can be used for attacks
against a company or an individual. More precisely, it wants to determine how
social engineering attacks can be driven against users of social networks,
based on data, which is legally accessible on social networks.

Of course, a distinction between employees of a certain company and private
users has to be made, as certain information can be harmless to an individual,
but a danger to a company~\cite{mitnick2003}. For example a phone number of a
private person is probably harmless, whilst the phone number of a certain
office inside a company could be used for a social engineering attack. In both
cases, however, a social engineer saves additional time and effort to get
specific information about his victim, if he can find the information needed on
a social network. Having to collect the information by hand does not only
require extra work, but it also includes the danger of being exposed. A social
engineer has to remain undiscovered during the attack, otherwise he would not
be able to carry out an attack~\cite{mitnick2003}.

Especially big corporations spend hundreds of thousands of dollars for the
security of their IT infrastructure. An attack on the IT layer would therefore
be often very laborious. Hence it is much simpler to bypass the security
mechanisms through social engineering, as it is very cheap and does not require
any superior technology~\cite{winkler1995}. Commonly, a skilled social engineer
does not require anything more than a telephone for an attack of such
nature~\cite{mitnick2003}.

Concretely, this study wants to answer the following research question:
How can data of individuals or companies automatically be extracted
from social networks and presented in a way that can be used for a social
engineering attack. In addition, countermeasures against automatic extraction
and social engineering attacks are going to be developed.

\section{Methodology}

In this work, a five-step approach was used to characterize useful information,
extraction, attacks and countermeasures.

\begin{enumerate}

\item Data was obtained from the chosen social network of several users. The
data was taken from well known people on the network, as well as less famous.
Furthermore, users who were very active such as less active users were
observed.

\item The data gathered in the first step was studied and it was determined
which kind of attacks could be realistic.

\item Three sample attacks were chosen, which took place in reality in the last
years. All of these attacks are described by Mitnick and
Simon~\cite{mitnick2003}. The chosen attacks were then transferred to the
prototype and the chosen social network.

\item The attacks were tested on sample profiles for their feasibility and
efficiency.

\item Countermeasures were elaborated to mitigate the risk of such attacks.

\end{enumerate}

\newpage
\section{Outline of the Thesis}

\let\oldchapterautorefname \chapterautorefname
\def\chapterautorefname{Chapter}

\noindent{\textbf{\autoref*{chap:introduction}}: \nameref{chap:introduction}}
\vspace{0.5em}\\
\noindent This chapter presents an overview of the thesis and introduces the
reader to the topic of new threats of social engineering by exploiting social
networks.

\noindent {\textbf{\autoref*{chap:relatedwork}}: \nameref{chap:relatedwork}}
\vspace{0.5em}\\
\noindent An outline of related researches in the field of social
engineering, the attacks and countermeasures will be presented. Common attacks
and how companies and individuals can protect themselves against such attacks
will also be showcased.

\noindent {\textbf{\autoref*{chap:attacks}}: \nameref{chap:attacks}}
\vspace{0.5em}\\
\noindent Three social engineering attacks, which happened in reality, were chosen
and are presented to the reader. They are analyzed in order to let the
prototype repeat the attacks.

\noindent {\textbf{\autoref*{chap:analysis}}: \nameref{chap:analysis}}
\vspace{0.5em}\\
\noindent The study takes a deeper look at the social network \Twitter. It will
show the drawbacks and risks of such social networks, the data which can be
extracted automatically and also, how sample attacks could look like. Finally the
countermeasures against such attacks are discussed.

\noindent {\textbf{\autoref*{chap:prototype}}: \nameref{chap:prototype}}
\vspace{0.5em}\\
\noindent In this chapter, a prototype is developed, which can automatically harvest data and
present it in a way that can be used for social engineering attacks.

\noindent {\textbf{\autoref*{chap:evaluation}}: \nameref{chap:evaluation}}
\vspace{0.5em}\\
\noindent This chapter examines whether the previously described attacks
can be achieved by using the prototype and therefore shows sample attacks. It
will analyze the attack and evaluate the risk against a company or an
individual.

\noindent {\textbf{\autoref*{chap:conclusion}}: \nameref{chap:conclusion}}
\vspace{0.5em}\\
\noindent The thesis finally concludes and draws together the main findings of the
study.

\def\chapterautorefname{\oldchapterautorefname}
\newpage

\section{Glossary}
\begin{multicols}{2}
\raggedcolumns

\begin{acronym}[XXXXX]

\acro{API}{An application programming interface (API) is an interface in computer science that defines ways to interact with services from libraries, the operating system or the web.}
\medskip
\acro{Atom}{The Atom Syndication Format is a format used for web feeds based on the XML Language.}
\medskip
\acro{GET}{A method for requesting a specified resource using the HTTP protocol.}
\medskip
\acro{JSON}{JavaScript Object Notation (JSON) is a computer data interchange format which uses the JavaScript syntax.}
\medskip
\acro{POST}{A method for transferring data to an identified resource using the HTTP protocol.}
\medskip
\acro{REST}{Representational State Transfer (REST) is a style for software architecture for distributed systems. The web is built on that style, for example.}

\columnbreak

\acro{RSS}{Really Simple Syndication (RSS) is a feed format to publish web feeds, like blogs, news, audio or video}
\medskip
\acro{URL}{A Uniform Resource Locator (URL) specifies how and where an identified resource is available and the methods retrieving it.}
\medskip
\acro{WHOIS}{A protocol usable for determining the registrant or owner of a domain name, IP adress or other Internet resources.}
\medskip
\acro{XML}{Extensible Markup Language (XML) is a general-purpose specification for creating custom markup languages.}

\end{acronym}

\end{multicols}
